\vspace{1.0cm} \Large \bf Integrating influenza antigenic dynamics with molecular evolution

### 2013-07-23 response.tex

Does a similar correlation analysis to the one shown in Fig. 4 find a relationship between estimated within-clade antigenic drift and excess P \& I mortality attributed to influenza?

\textbf{We have begun to extend the correlation analysis between within-clade antigenic drift and multiple measures of influenza incidence, including ILI, virus isolation, excess P\&I mortality and influenza-associated hospitalization rates. This analysis is also attempting to look at a finer temporal scale, rather than the course-grained 1-year bins in the current study and attempting to better match geographic bounds in antigenic and incidence data.  This said, we believe that further extending the incidence correlation requires a separate publication to  describe adequately, and is beyond the scope of the current analysis.}

The ``uninformed'' prior used in Table 1 is not described in detail, but it actually hides a subtle issue. For an unbounded variable like the x- and y- positions of a mapped isolate, the ``uninformed'' prior cannot simply be uniform on $R^2$, but rather is uniform on some compact subset of $R^2$. In other words, the authors had to assume a prior with a maximum possible x- and y-position. They don't describe what maximum bounds went into their ``uniformed'' prior, but this is well-known to influence the results of Bayesian analyses in general. This is a relatively minor issue -- authors should simply show how the form of his uniformed prior influences the resulting maps.

\textbf{We apologize for the lack of clarity.  The `Uninformed location prior' in Table 1 was described in Materials and methods as a diffuse normal prior (equation 10) with each coordinate having a variance of 10000.  We've made this clearer in the test.  We chose this variance based on prior knowledge of the scale of antigenic maps, i.e. around 50 units in breadth for A/H3N2 from Smith et al.\ 2004.  In general, our use of diffuse priors limits their influence on the resulting antigenic maps.  In this case, choosing a variance of 1000 gives very similar results.}

### 2013-07-23 flux.tex
Fix #29 - use of "uninformed" prior.

We begin with Bayesian analogs of the models used by Smith et al.\ \cite{Smith04}, in which viruses and sera are represented as $N$-dimensional locations, expected titer is relative to the maximum titer of a particular ferret serum and virus and serum locations follow a diffuse prior.
After comparing models of differing dimensions, Smith et al.\ \cite{Smith04} arrive at a 2D model as the preferred model for their data.
Smith et al.\ \cite{Smith04} implement a form of MDS, seeking to optimize virus and serum locations such that the sum of squared errors between expected and observed titers is minimized (eq.\ \ref{mds} of Materials and methods).
Here, in implementing BMDS, we provide a likelihood function for the probability of observing HI data given virus and serum locations (eq.\ \ref{bmds} of Materials and methods) and seek to estimate model parameters through Bayesian inference using Markov chain Monte Carlo (MCMC).
However, the basic antigenic model describing drop in HI titer as proportional to Euclidean distance between virus and serum locations is identical between these methods.

>>>>

We begin with Bayesian analogs of the models used by Smith et al.\ \cite{Smith04}, in which viruses and sera are represented as $N$-dimensional locations.
In this case, `serum effects' are fixed to the maximum titers exhibited by particular ferret sera, and virus and serum locations follow an uninformative diffuse normal prior as described in the `Antigenic cartography' section of Materials and methods.
After comparing models of differing dimensions, Smith et al.\ \cite{Smith04} arrive at a 2D model as the preferred model for their data.
Smith et al.\ \cite{Smith04} implement a form of MDS, seeking to optimize virus and serum locations such that the sum of squared errors between expected and observed titers is minimized (eq.\ \ref{mds} of Materials and methods).
Here, in implementing BMDS, we provide a likelihood function for the probability of observing HI data given virus and serum locations (eq.\ \ref{bmds} of Materials and methods) and seek to estimate model parameters through Bayesian inference using Markov chain Monte Carlo (MCMC).
However, the basic antigenic model describing drop in HI titer as proportional to Euclidean distance between virus and serum locations is identical between these methods.
